\chapter{Endogenous configuration space}
\label{ch:endogen}
\section{Basic concepts}
Let us consider a mobile robot --- a mobile platform with a manipulator. Let $q = (w, \dot w)^T \in \mathbb{R}^{2n}$ denote the state of the platform and $x \in \mathbb{R}^p$ denote the configuration of the manipulator. The platform is actuated by vector $u \in \mathbb{R}^m$. We can define a vector $y \in \mathbb{R}^r$ which is the result of the output function $k: \mathbb{R}^{2n} \times \mathbb{R}^p \rightarrow \mathbb{R}^r$. 

Suppose the dynamics of the platform are defined by the equation obtained using Euler-Lagrange formalism:
\begin{equation}
Q(w)\ddot w + C(w, \dot{w})\dot{w}+D(w)=F(w, \dot w)+B(w)u,
\end{equation}
where $Q(w)$ is the inertia matrix, $C(w, \dot w)$ is the centripetal and Coriolis forces matrix,
$D(w)$ is the potential forces vector, $F(w, \dot w)$ is the vector of traction forces and $u$ is the
control forces vector applied to the system having been transformed by matrix $B(w)$.
This system may be transformed into the form $\dot q = f(q) + G(q)u$ using
\begin{equation}
\begin{aligned}
f(q)&=\begin{bmatrix}
\dot{w}\\
Q^{-1}(w)\left(F(w, \dot w)-C(w, \dot{w})\dot{w}-D(w)\right)
\end{bmatrix}, & G(q)&=\begin{bmatrix}
0_{n\times m}\\
Q^{-1}B
\end{bmatrix}
\end{aligned}
\end{equation}

Adding to the above the output function depending on the platform state and the configuration of the manipulator gives the following affine control system with inputs
$u \in \mathbb{R}^m$, $x \in \mathbb{R}^p$ and output function $k$.
\begin{equation}
\begin{cases}
\begin{aligned}
\label{eq:control_sys}
\dot q &= f(q) + G(q)u,\\
y &= k(q, x).
\end{aligned}
\end{cases}
\end{equation}

\subsection{Endogenous configuration}
Define a pair $(u(\cdot), x)$, which consists of control inputs mentioned above.
Such elements belong to the endogenous configuration space
$\mathcal{X} = L_m^2[0, T] \times \mathbb{R}^p$ of the mobile manipulator.
The inner product in $\mathcal{X}$ is as follows \cite{ecs_ijc}:
\begin{equation}
\langle (u_1(\cdot), x_1), (u_2(\cdot), x_2) \rangle = \int_0^T u_1^T(t) u_2(t) \ud t + x_1^T x_2.
\end{equation}

The function
$K_{q_0, T}: \mathcal{X} \rightarrow \mathbb{R}^r$
allow to determine the state of the mobile manipulator at time $T>0$ given
endogenous configuration $(u(\cdot), x)$ and initial state $q_0$:
\begin{equation}
K_{q_0, T}((u(\cdot), x)) = k(\phi_{q_0, T}(u(\cdot)), x),
\end{equation}
where $\phi_{q_0, T}(u(\cdot))$ denotes the flow of the system \eqref{eq:control_sys}
caused by input $u(\cdot)$. The above will be called the end-point map.

With the above definitions the analytic Jacobian of $K_{q_0, T}$ can be defined at $(u(\cdot), x) \in \mathcal{X}$ \cite{ratajczak2013multiple}:
\begin{align}
\label{eq:jacobian}
J_{q_0, T}(u(\cdot), x)(v(\cdot), w) &= \left.\frac{\ud}{\ud \alpha}\right|_{\alpha=0} K_{q_0, T}((u(\cdot)+\alpha v(\cdot), x+\alpha w))\\
 &= 
 C(T,x)\int_0^T \Phi(T,s)B(s)v(s)\ud s + D(T,x)w,
\end{align}
where $(v(\cdot), w)\in \mathcal{X}$ and $\Phi(T,s)$ is the fundamental matrix of the following system:
\begin{equation}
\begin{aligned}
\label{eq:abxi}
\dot \xi &= A(t)\xi + B(t) v \\
\eta &= C(t, x)\xi + D(t, x).
\end{aligned}
\end{equation}
The matrices in \eqref{eq:abxi} are as follows \cite{tchon2004acceleration}
\begin{equation}
\begin{aligned}
A(t) &= \frac{\partial (f(q(t))+G(q(t))u(t)}{\partial q}, & C(t, x) &= \frac{\partial k(q(t), x)}{\partial q},\\
B(t) &= G(q(t)), & D(t, x) &= \frac{\partial k(q(t), x)}{\partial x}.
\end{aligned}
\end{equation}

\section{Motion planning algorithm}
Given the end-point map $K_{q_0, T}$, the motion planning problem is equivalent to finding an endogenous
configuration $(u_d(\cdot), x_d)$, such that $K_{q_0, T}((u(\cdot), x))=y_d$. In order to find the
solution a Jacobian algorithm may be employed, analogically to the algorithms
for inverse kinematics problem for manipulators. 

Let us choose the initial configuration $(u_0(\cdot), x_0)$ and
define a curve in $\mathcal{X}$ $(u_\theta(\cdot), x_\theta)$, $\theta \in \mathbb{R}$.
Computing the error along $(u_\theta(\cdot), x_\theta)$ we obtain
\begin{equation}
\label{eq:endogen_err}
e(\theta)=K_{q_0, T}(u_\theta(\cdot), x_\theta)-y_d.
\end{equation}
Since we expect the error to stabilize at $0$ exponentially fast with decay rate
$\gamma$ we get the formula
\begin{equation}
\label{eq:endogen_decay}
\frac{\ud e(\theta)}{\ud \theta}=-\gamma e(\theta)
\end{equation}.
Combining \eqref{eq:endogen_err} and \eqref{eq:endogen_decay} we acquire the Ważewski-Davidenko equation
\begin{equation}
J_{q_0, T}(u_\theta(\cdot), x)\dfrac{\ud}{\ud \theta}\begin{pmatrix}
u_\theta(\cdot)\\ x_\theta
\end{pmatrix}=-\gamma e(\theta).
\end{equation}

Let $J^\#_{q_0, T}(u_\theta(\cdot), x)$ denote the right Moore-Penrose inverse
of the Jacobian \eqref{eq:jacobian}. Applying it to the above equation we get a dynamic system
\begin{equation}
\label{eq:endogen_alg}
\dfrac{\ud}{\ud \theta}\begin{pmatrix}
u_\theta(\cdot)\\ x_\theta
\end{pmatrix}=-\gamma J^\#_{q_0, T}(u_\theta(\cdot), x) e(\theta),
\end{equation}
which defines the solution of the formulated motion planning task in the following way:
\begin{equation}
\begin{pmatrix}
u_d(\cdot)\\ x_d
\end{pmatrix}=\lim_{\theta\rightarrow\infty}\begin{pmatrix}
u_\theta(\cdot)\\ x_\theta
\end{pmatrix}.
\end{equation}

\section{Numerical computations}
The dimension of $L^2[0,T]$ space is infinite. Therefore we need an infinite number of parameters to define an element from this space. This is impossible to achieve in calculations made on a computer. The solution is to limit the bandwidth of considered functions. 

It is possible to choose a finite-dimensional orthogonal base $\Psi = [ \psi_1, \psi_2, \dots, \psi_s ]$ spanning the limited-band control functions space. 
Using such approach a function $u_i$ is represented by a vector 
$\lambda_i = [\lambda_{i1}, \lambda_{i2}, \dots, \lambda_{is}]^T$ 
which satisfies the formula $u_i = \sum_{k=1}^s \psi_k \lambda_{ik}$.

In order to define a vector $u=[u_1, u_2, \dots, u_m]^T$ a matrix $P(t)$ should be defined, such that $u_\lambda(t)=P(t)\lambda$. This means $P(t)$ is a block diagonal matrix 
\begin{equation}
P(t)=\begin{bmatrix}
\Psi & 0 & \cdots & 0\\
0 & \Psi &  & \vdots\\
\vdots &  & \ddots & 0 \\
0 &  \cdots & 0 & \Psi
\end{bmatrix}
\end{equation}

The above formulae let us define a new band-limited Jacobian as
\begin{equation}
\hat J_{q_0, T}(\lambda, x)=\begin{bmatrix}C_\lambda(T)\int_0^T \Phi_\lambda(T,s)B(s)_\lambda P(s)\ud s & D_\lambda(T,x)\end{bmatrix}
\end{equation}

Assuming the above the Jacobian $\hat J_{q_0, T}(\lambda, x)$
may be computed alternatively as\\ $\begin{bmatrix}
C\cdot\Xi(T)& D
\end{bmatrix}$, where $\Xi$ is defined by  
\begin{equation}
\dot \Xi = A_\lambda(t)\Xi +B_\lambda(t)P(t)
\end{equation}
with the initial condition $\Xi(0)=0$ %TODO citation needed.
This allows to easily calculate the Jacobian simultaneously with the simulation
of the system by extending the ODE function.

In order to acquire the Moore-Penrose inverse of the Jacobian we can simply use the formula
\begin{equation}
\hat J^\#_{q_0, T}(\lambda, x)=\hat J^T_{q_0, T}(\lambda, x)\left(\hat J_{q_0, T}(\lambda, x)\hat J^T_{q_0, T}(\lambda, x)\right)^{-1}.
\end{equation}

The problem which may occur here is the singularity of the dexterity matrix\\
$\hat J_{q_0, T}(\lambda, x)\hat J^T_{q_0, T}(\lambda, x)$. The inverse can be regularised by adding a small value to the diagonal of the matrix being inverted in the following way:
\begin{equation}
\hat J^\#_{q_0, T}(\lambda, x)=\hat J^T_{q_0, T}(\lambda, x)\left(\hat J_{q_0, T}(\lambda, x)\hat J^T_{q_0, T}(\lambda, x)+\eta I\right)^{-1}
\end{equation}
The regularised version of the inverse Jacobian should be used only if the determinant of the dexterity matrix is below the assumed threshold.

Another limitation met in numerical approach to the problem is the necessity to discretise the variable $\theta$ in \eqref{eq:endogen_alg}.
This may be resolved by applying the first-order Euler method:
\begin{equation}
\begin{pmatrix}
\lambda_{k+1}\\
x_{k+1}
\end{pmatrix}  =\begin{pmatrix}\lambda_{k}\\
x_{k}
\end{pmatrix} - \gamma \hat J^\#_{q_0, T}(\lambda_k, x_k)e(k).
\end{equation}